## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}
a <- 10

if(a<=5){
  print(a)
}else{
  print("It is more than 5")
}
```

We can also add else if statements in the middle if we need more options.

```{r}
a <- 15
if(a<=5){
  print("Small")
}else if(a > 5 & a <= 10){
  print("Medium")
}else if(a > 10 | a == 7){
  print("Large")
}else{
  print("Huge")
}


```

Neat! We can also nest conditionals just like we do in Python.

```{r}
a <- 5
if(a/2 != 0){
  if(a ==5){
    print("a is 5")
  }
}

    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}
for (value in 1:5) {
  #paste is a concatenation function
  print(paste(value, "is greater than", value - 1))
}

```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}
a_vector <- c(1,1,1,1,1,1,1)

# Change the first, third, sixth elements to 6
for (i in c(1,3,6)) {
  a_vector[i] <- 6
  
}
a_vector

```

We can loop over more than just lists:

```{r}
mat <- matrix(1, nrow = 2, ncol = 2)
for (row in (1:2)) {
  for (col in (1:2)) {
    mat[row, col] <- 0
    print(mat)
    
  }
  
}

```

Like Python, we have while loops in R:

```{r}
b <- 0
while(b < 10){
  #increase b by 2
  b <- b + 2
  print(b)
}
```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
```

An example:

```{r}
#Start with a vector of tempertures
temp_test <- c (19.56785, 26.98762, 32.86543, 8.65432, 21.54329,18.76420, 
                23.90673 )

#Then we want a function that converts temp to C to F
c_to_f <- function(dat){
  temp_farh <- dat*(9/5) + 32
  return(temp_farh)
}
print(c_to_f(26))
```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}
data_fahr <- c_to_f(temp_test)
data_fahr

```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}
test_mat <- matrix(c(1,1,1,2,2,2),nrow = 2)
test_mat
# 1 is for row and 2 for columns
apply(test_mat, 2, sum)

```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}
test_temp_mat <- matrix(c(19.56785, 26.98762, 32.86543, 8.65432, 21.54329,18.76420,               23.90673 ),nrow = 2)
apply(test_temp_mat, 1, c_to_f)

```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
    1.#Third Angle of a Triangle

    	other_angle <- function(a, b){
     	return (180-(a+b))
    }

    2.#Get Nth Even Number

    	nth_even <- function(n) {
      return (2* n-2)
    }

    3.#You Can't Code Under Pressure #1

    	double_integer <- function(i) {
    return (i * 2)
    }

    4.#Is he gonna survive?

    	hero <- function(bullets, dragons) {
     return( bullets >= 2* dragons)
      }

    5.#Remove First and Last Character

    	library(stringr)
    removeChar <- function(str){
      substr(str, 2, nchar(str) - 1)
    }


    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
    num_test <- c(0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20)

    number_fn <- function(num){
      if(num %%3 == 0){
        return("fizz")
      } else if (num %% 5 == 0){
        return("buzz")
        
      }else if(num %% 15 == 0){
        return("fizzbuzz")
      }else{
        print(num)
      }
    }

    #print(number_fn(15))
    print(number_fn(5))
    #print(number_fn(9))




    ```

    ```{r}


    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.

    marioStairs <- function(height) {
      for (i in 1:height) {
        cat(rep("X", i), "\n")
      }
    }
    marioStairs(4)

    ```

```{r}
marioStairs <- function(height) {
  stairs <- sapply(1:height, function(i) {
    paste(rep("X", i), collapse = "")
  })
  cat(paste(stairs, collapse = "\n"))
}
marioStairs(4)
```
