## **R Libraries: ggplot2**

The R library `ggplot2` is a super useful tool we can use to create visualizations and explore our data.

```{bash}
install.packages("tidyverse")
library("tidyverse")
```

More things we can do to get more information on R and what we're doing:

```{bash}
help()
```

We'll need a few more packages for our exploration usage, too. We can install them all at once by using the `c()` function, which creates a list. We'll open these packages with the `library()` function later.

```{bash}
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

Built in to our tidyverse are several data sets that we can play around with. In the below code block, we create a variable `myplot` and save the `mpg` database to it.

```{r}
my_plot <- mpg
mpg

```

**Something you should be aware of and have open (perhaps on a second monitor, or you can print them out) is the ggplot2 cheat sheet!** <https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf>

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))

```

```{r}
mtcars

```

```{r}
ggplot(data = mpg)+geom_point(mapping = aes(x = displ, y = hwy, color=class))
#color = displ < 5
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy ,color = rgb(0.3,1,0.1))) +
  facet_wrap(~ class, nrow = 3)
#color = class


```

```{r}
ggplot(data = mpg)+ geom_point(mapping = aes(x = displ, y = hwy, color = drv))+
  facet_grid(drv ~ .)

```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv)) +
geom_point(mapping = aes(color = class)) +
geom_smooth(se = FALSE)  

```

```{r}

```

```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut))  

```

Whoa, where'd count come from? **Count is not a variable or attribute of the variable diamonds!** Count comes automatically with a lot of graphs as an attribute. You can verify what I'm saying with `?diamonds` in the console.

Any attribute of a data set that is algorithmically calculated is called a *stat*, which is short for a statistical transformation. Many of the **geom** functions have stats built in, and many stats display geoms. For example, the above code block used a geom, but this one uses a stat, and results in an identical chart:

```{r}
ggplot(data = diamonds)+
  stat_count(mapping = aes(x = cut))

```

Let's add color here!

```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut, fill = cut))

```

What if we add to a different variable than cut?

```{r}
pie = ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut, fill = clarity), position = 'fill')+
  theme(aspect.ratio = 1)+
  labs(x = NULL, y = NULL)

pie + coord_polar()

```

**For your consideration...** Is the above a good graph?

```{r}
alt- becomes <- (option- on mac)
```

## Homework

```{r}
# Find a builtin data set for students to explore; use their cheat sheet from stats to label continuous and discrete data; use graphs to tell me things about 3 datasets.
view(nycflights13::flights)
```

```{r}


# plot the number of flights by month
ggplot(data = flights, aes(x = month)) +
  geom_bar(fill = "red") +
  labs(title = "Number of Flights by Month", x = "Month", y = "Number of Flights")


```

```{r}
# plot the average delay by carrier
ggplot(data = flights)+
geom_point(mapping = aes(x = carrier, y = arr_delay, color= origin)) 
 
```

```         
```

```{r}



```

```{r}
# plot the relationship between air time and distance
ggplot(data = flights, mapping = aes(x = distance, y = air_time)) +
  geom_point(mapping = aes(alpha = 0.5, color = origin))
   +
  labs(title = "Relationship Between Air Time and Distance", x = "Distance (miles)", y = "Air Time (minutes)")



```

```{r}
my_plot <- beaver1
beaver1 
view(beaver1)
```

```{r}
ggplot(data = beaver1)+
geom_point(mapping = aes(x=time, y= temp, color = activ ))
```

```{r}
ggplot(data = beaver1)+
  geom_bar(mapping = aes(x = day)) 

```

```{r}
my_plot <- trees
trees
view(trees)

```

```{r}
ggplot(data = trees)+
geom_point(mapping = aes(x=Girth, y= Height))

```

```{r}
ggplot(data = trees)+ geom_point(mapping = aes(x = Volume, y = Girth,
color = Height))+
  facet_grid(Height ~ .)
```

```{r}
my_plot <- HairEyeColor
HairEyeColor
view(HairEyeColor)

```

```{r}
data(HairEyeColor)
df <- melt(HairEyeColor)

```

```{r}
str(HairEyeColor)
```

```{r}
# Convert table to data frame
df <- as.data.frame.table(HairEyeColor)

# Create scatter plot
ggplot(df, aes(x = Eye, y = Hair, color = Sex, size = Freq)) +
  geom_point() +
  labs(x = "Eye Color", y = "Hair Color", color = "Sex", size = "Frequency")



```

```{r}


# Convert table to data.frame
HairEyeColor_df <- as.data.frame(HairEyeColor)

# Create bar plot
ggplot(HairEyeColor_df, aes(x = Sex, y = Freq, fill = Hair)) +
  geom_col(position = "dodge") +
  labs(title = "Hair and Eye Color by Sex",
       x = "Sex", y = "Frequency") +
  theme_minimal()






```

```{r}
my_plot <- ChickWeight
view(ChickWeight)


```

```{r}
ggplot(data = ChickWeight) +
  geom_point(mapping = aes(x =Chick , y=weight  ,color = Diet)) +
  facet_grid(Diet ~ .)
```

```{r}
ggplot(data = ChickWeight)+
  geom_point(mapping = aes(x = Time, y = weight, color = Diet))


```

```{r}
d <- ggplot(ChickWeight, aes(weight))
d + geom_histogram(bindwidth = 5, color = "black", fill = "red")+
labs(title = "Chik wrt to weight", x ="Weight", y = "No.of chicks")+
theme_classic()  


```

```{r}
p1 <-ggplot(mtcars, aes(factor(cyl),mpg))
p1 + geom_boxplot()
p1 +geom_boxplot() + coord_flip()
p1 +geom_boxplot(aes(fill = factor(cyl)))

```

```{r}



```
